---
title: "Boston Crime Data Analysis"
output:
  pdf_document: default
  html_notebook: default
---

 

```{r}
library(tidyverse)
library(ggplot2)
library(lubridate)
library(ggrepel)
library(forcats)
library(scales)
library(dplyr)
library(sf)
library(mapview)
library(ggmap)
library(ggpubr)
library(deforestable)
```

```{r}
#importing the data
boston_crime_df <- read.csv("C:\\Users\\17579\\Desktop\\Boston_Crime_Data.csv")
```

```{r}
str(boston_crime_df)
summary(boston_crime_df)
```


```{r}
boston_crime_df$INCIDENT_NUMBER <- as.factor(boston_crime_df$INCIDENT_NUMBER)
boston_crime_df$OFFENSE_CODE_GROUP <- as.factor(boston_crime_df$OFFENSE_CODE_GROUP)
boston_crime_df$OFFENSE_DESCRIPTION <- as.factor(boston_crime_df$OFFENSE_DESCRIPTION)
boston_crime_df$DISTRICT <- as.factor(boston_crime_df$DISTRICT)
boston_crime_df$SHOOTING <- as.factor(boston_crime_df$SHOOTING)
boston_crime_df$OCCURRED_ON_DATE <- as.Date(boston_crime_df$OCCURRED_ON_DATE, format = "%m/%d/%y")
boston_crime_df$DAY_OF_WEEK <- as.factor(boston_crime_df$DAY_OF_WEEK)
boston_crime_df$UCR_PART <- as.factor(boston_crime_df$UCR_PART)
boston_crime_df$STREET <- as.factor(boston_crime_df$STREET)
boston_crime_df$Location <- as.factor(boston_crime_df$Location)
boston_crime_df$YEAR <- as.factor(boston_crime_df$YEAR)

```

### Boston's Crime Trend

```{r}
# Yearly crime trend in Boston, MA.

boston_crime_df %>%
  select(YEAR) %>%
  group_by(YEAR) %>%
  summarise(count_k = n()) %>%
  ggplot()+
  geom_bar(fill = 'blue', stat = 'identity', aes(x = YEAR, y = count_k))+
  ggtitle("Crime Reported vs Year (Boston, MA)")+
  ylab("No. of crimes reported")
```
Fewer records exist for years 2015 and 2020 as compared to rest of the years.
 
```{r}
boston_crime_df %>%
  group_by(YEAR, MONTH) %>%
  summarise(
    count_crimes = n()
  ) %>%
  summarise(
    average_monthly_crime = mean(count_crimes)
  )
```
Similar monthly crime rate over the years.
```{r}
boston_crime_df %>%
  group_by(YEAR, MONTH) %>%
  summarise(
    count_crimes = n()
  ) %>%
  ggplot(aes(x = MONTH, y = count_crimes))+
  geom_line(color = "blue")+
  geom_point(color = "red") +
  scale_x_discrete(limits = month.abb)+
  facet_grid(rows = vars(YEAR))+theme_bw()
```
Recording starts from June 2015 and ends at April 2020.
Monthly crimes follows similar pattern each year.

```{r}
boston_crime_df %>%
  mutate(weekend = if_else(DAY_OF_WEEK == "Saturday" | DAY_OF_WEEK == "Sunday", "Yes", "No")) %>%
  group_by(MONTH, weekend) %>%
  summarise(
    no_of_days = n()
  ) %>%
  ggplot(aes(x = weekend, y = no_of_days))+
  geom_boxplot()+
  ggtitle("Crime Reported : Weekdays vs Weekends (Boston, MA)")+
  ylab("No. of crimes reported")+
  xlab("Weekend")
```

```{r}
boston_crime_df %>%
  group_by(MONTH, DAY_OF_WEEK) %>%
  summarise(
    no_of_crimes = n()
  ) %>%
  ggplot(aes(x = DAY_OF_WEEK, y = no_of_crimes))+
  geom_boxplot()
  
```
### Crime Pattern in Boston's Districts and Streets.

```{r}
ggplot(boston_crime_df, aes(x=fct_infreq(DISTRICT)))+
  geom_bar(stat = "count")
```

```{r}
b1 <- boston_crime_df %>%
  filter(DISTRICT != "") %>%
  group_by(DISTRICT, STREET) %>%
  summarise(
    countk = n()
  ) %>%
  arrange(DISTRICT, desc(countk))

b1
```

```{r}

b1 %>%
  ggplot(aes(x = countk))+
  geom_histogram(bins = 50, fill = "red", color = "black")+
  scale_x_continuous(limits = c(0,1000))+
  scale_y_continuous(limits = c(0,150))+
  facet_wrap(~ DISTRICT, ncol = 3)+
  theme_get()+
  ggtitle("Crime Reported per Streets in Boston, MA")+
  ylab("No. of Streets")+
  xlab("No. of Reported Crimes")
```

Majority of the streets had reported less than 250 crimes during the period 2015 - 2020.


```{r}
b2 <- b1 %>%
  filter(DISTRICT != "External") %>%
  mutate(crime_rate_indicator = case_when(
    countk > 800 ~ "high",
    countk > 200 & countk <= 800 ~ "mid",
    countk >0 & countk <= 200 ~ "low"
  ))

b2 %>%
  ggplot(aes(x=fct_infreq(DISTRICT), y = countk, fill = crime_rate_indicator))+
  geom_bar(stat = "identity")+
  scale_fill_discrete(labels = c("Streets with > 800 crimes", "Streets with <= 800 crimes", "Streets with less than 200 crimes"))+
  ggtitle("Crime Reported by Streets in Boston, MA")+
  ylab("No. of Streets")+
  xlab("Districts")
```

```{r}
ggplot(b2, (aes(x=fct_infreq(DISTRICT), y = countk, fill = crime_rate_indicator)))+
geom_bar(position = "fill", stat = "identity")+
ggtitle("Graph showing Crime Rate Proportions in Streets")+
theme(plot.title = element_text(hjust = 0.5))+
  ylab("No. of Streets")+
  xlab("Districts")
```
District D4 has the most number of streets with more than 800 reported crimes over the course of four years (2015 - 2020).

### UCR Part One Crimes in Boston, MA

```{r}
boston_crime_df %>%
  filter(UCR_PART == 'Part One') %>%
  group_by(DISTRICT) %>%
  summarise(
    count_of_partone = n()
  ) %>%
  arrange(desc(count_of_partone))


  
```


```{r}
#most frequent Part One crimes by street

b3 <- boston_crime_df %>%
  filter(UCR_PART == "Part One") %>%
  filter(DISTRICT != "") %>%
  group_by(DISTRICT, OFFENSE_CODE_GROUP) %>%
  summarise(
    countofcrimes = n()
  )%>%
  arrange(DISTRICT, desc(countofcrimes))
  
b3
```

```{r}
b3 %>%
  ggplot(aes(x = OFFENSE_CODE_GROUP, y = countofcrimes))+
  geom_bar(stat = "identity")+
  coord_flip()+
  facet_wrap(~DISTRICT)
```

```{r}
# pie chart showing the proportion of part one crimes in Boston from 2016 - 2020.

b4 <- b3 %>% 
  group_by(OFFENSE_CODE_GROUP) %>%
  arrange(desc(OFFENSE_CODE_GROUP)) %>%
  mutate(prop = countofcrimes / sum(b3$countofcrimes) *100) %>%
  mutate(ypos = cumsum(prop)- 1*prop )

ggplot(b4, aes(x="", y=countofcrimes, fill=OFFENSE_CODE_GROUP)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0)+
  theme_void()
```


```{r}
boston_crime_df %>%
  filter(UCR_PART != "" & UCR_PART != "Other")%>%
  ggplot(aes(x = DISTRICT, fill = UCR_PART))+
  geom_bar(position = "fill", stat = "count")+
  ggtitle("Street Level Crime Reported by District in Boston, MA")+
  ylab("Proportions of Streets")+
  xlab("Districts")
```

### Finding most dangeorus streets in each district.

```{r}
street_dangerous <-boston_crime_df %>%
  group_by(DISTRICT, STREET, UCR_PART) %>%
  summarise(
    street_crime = n()
  ) %>%
  filter(street_crime > 800 & UCR_PART == "Part One")
street_dangerous
```


```{r}
boston_crime_df %>%
  group_by(DISTRICT, STREET, UCR_PART) %>%
  summarise(
    street_crime = n()
  ) %>%
  filter(street_crime > 800 & UCR_PART == "Part Two")
```

```{r}
shoot <- boston_crime_df %>%
  filter(SHOOTING == "1" | SHOOTING == "Y") %>%
  group_by(DISTRICT, STREET) %>%
  summarise(
    shoot_count = n()
  ) %>%
  filter(shoot_count > 20) %>%
  arrange(DISTRICT, desc(shoot_count))

shoot
```

```{r}
shoot %>%
  filter(DISTRICT == "B2") %>%
  filter(shoot_count > 20) %>%
  ggplot(aes(x = STREET, y = shoot_count))+
  geom_bar(stat = "identity")+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))
```


```{r}
shoot %>%
  filter(DISTRICT == "C11") %>%
  filter(shoot_count > 20) %>%
  ggplot(aes(x = STREET, y = shoot_count))+
  geom_bar(stat = "identity")
```

```{r}

boston_crime_df %>%
  filter(SHOOTING == "1" | SHOOTING == "Y") %>%
  group_by(DISTRICT, MONTH) %>%
  summarise(
    monthly_shooting = n()
  ) %>%
  ggplot(aes(x = MONTH, y = monthly_shooting, col = DISTRICT))+
  geom_line()+
  scale_x_discrete(limits = month.abb)+
  ggtitle("Monthly Shooting Frequencies during 2015 - 2020")
```

